\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{A Radial Time Projection Chamber for $\alpha$ detection in CLAS at JLab}

### 2017-05-23 eg6-letters/NIM/EG6_RTPC_NIM.tex
Small edits

The GEM technology has been chosen for this RTPC for its flexibility, 
allowing a curved amplification surface. Also, GEMs are known to rarely create 
sparks{\color{blue} \cite{GEM_ref}}, which is important when trying to detect 
highly ionizing slow nuclei that deposit large amount of energy. The GEMs for 
this RTPC are made from a Kapton insulator layer, {\color{blue}50~$\mu$m 
thick}, sandwiched between two 5~$\mu$m copper layers. The mesh of each GEM 
layer is chemically etched with 50~$\mu$m diameter holes with double-conical 
shapes as illustrated in figure~\ref{fig:GEMs}. A potential difference of 275 V 
is applied between the two 
copper layers of the GEM, creating a very strong electric field in each hole 
leading to high ionization and therefore amplification of the incoming 
electrons. A high potential of about 1600 V/cm is set between 
the GEMs to push the electrons toward the next GEM and eventually the readout 
pads. \\

The drift gas used in the experiment is a 80-20\% neon-DME mixtur, this choice 
has been made in order to balance the good drift properties with a reasonnable
Lorentz angle. Calculations using the MAGBOLTZ program~\cite{MAGBOLTZ} 
showed that with the perpendicular 5~T field, we would have a Lorentz angle of 
about {\bf \color{red} (To be added) } with this mixture. As can be seen, the
Lorentz angle is already significant, a larger angle would increase significantly
the length of the drift paths as the electron would spiral around the beam line.
{\bf \color{red} (To be continued commenting on higher/lower proportion of gas 
and other gas) }

One of the important problematic to develop the radial TPC was to obtain a 
good support structure of the GEM foils, allowing an easy installation of the 
GEM in good conditions and with high precision. At the same time, we wanted to 
keep the material budget small in the forward region where we detect other 
particles in subsequent detectors. We successfuly realized these prerequisites 
by using fiber glass rings glued to each ends of the GEM foils to form self 
supporting cylinders that could be installed independently in the RTPC after 
all gluing and soldering operations.  Because the rigidity of the high 
rigidity of the GEM foil, the strcture was selfsupporting and only the 
backward part of the cylinder was fixed to the main mechanical structure. This 
design only left a light fiberglass ring in the forward region, reducing to a 
minimum secondary interactions.

@@

The drift path is the trajectory followed by the electrons released through 
ionization in the gas. Software exists to calculate the drift paths, in particular
MAGBOLTZ~\cite{MAGBOLTZ}, but it requires knowledge of the detector's 
geometry, gas mixture composition, and of course the electric and magnetic 
fields over the whole volume of the detector. We used such calculation as a 
first calibration, but, 
as can be seen with the drift speed, gas composition is far from stable in the 
chamber. Moreover, the 4 $\mu$m foil used as a cathode is easily deformed, such
that we expect the geometry accuracy to be around the millimeter precision. 
These problems, already 
encountered for the BoNuS RTPC calibration~\cite{BONUS-NIM}, motivated the 
acquisition of specific calibration runs. These are taken with a lower energy 
electron beam (1.20 and 1.27 GeV) to enhance the cross section of the elastic 
process ($e^{4}$He$\rightarrow e^{4}$He)
{\bf \color{red} Specify luminosity ?}. In this process, the measurement of
electron kinematics allows to calculate the Helium nucleus kinematic. It is
by comparing this calculated kinematic to the measured one that we fine tune
the drift paths independently of our knowledge of the exact conditions in the 
chamber.

@@

The second method is to compare the experimental ADCs to the energy deposited in 
GEANT4 by similar simulated tracks (using the same elastic events than for the drift 
paths calibration). This requires a very good GEANT4 simulation 
including drift paths, but also the spread of the charges along the path
before reaching the pad, so that the simulated hits match the experimental 
ones. Moreover, the simulation has to match the data acquisition (DAQ) features 
that can lead to cutting out hits. After setting the simulation properly, we 
compared simulation to experiment on an event by event basis as shown in 
figure~\ref{fig:EVENT_adc_tdc}. In this step, the gain for each pad is defined 
as the ratio of the measured ADCs to the simulated deposited energy.  
{\color{blue} Then, the gain of each readout pad is corrected by a factor, that 
is defined as the ratio of the ADC value of each pad to the ADCs of the other 
pads inside each track to make sure that the recorded ADCs by a given pad are 
similar to the recorded ADCs recorded by other pads in the same track.}

{\bf \color{red} Mohammad: Detail are missing here.}
Results of this second method are shown in Figure~\ref{fig:dedx_p_exp_2nd},
from these we concluded that the gains of the 
second method match best the theoretical lines.

>>>>

The GEM technology has been chosen for this RTPC for its flexibility, 
allowing a curved amplification surface. Also, GEMs are known to rarely create 
sparks~\cite{GEM_ref}, which is important when trying to detect 
highly ionizing slow nuclei that deposit large amount of energy. The GEMs for 
this RTPC are made from a Kapton insulator layer, 50~$\mu$m 
thick, sandwiched between two 5~$\mu$m copper layers. The mesh of each GEM 
layer is chemically etched with 50~$\mu$m diameter holes with double-conical 
shapes as illustrated in figure~\ref{fig:GEMs}. A potential difference of 275 V 
is applied between the two 
copper layers of the GEM, creating a very strong electric field in each hole 
leading to high ionization and therefore amplification of the incoming 
electrons. A high potential of about 1600 V/cm is set between 
the GEMs to push the electrons toward the next GEM and eventually the readout 
pads. \\

The drift gas used in the experiment is a 80-20\% neon-DME mixture, this choice 
has been made in order to balance the energy deposit with a reasonnable
Lorentz angle. Calculations using the MAGBOLTZ program~\cite{MAGBOLTZ} 
showed that with the perpendicular 5~T field, we would have a Lorentz angle of 
about 23$^\circ$ with this mixture. This is already significant and a heavier
gas would have larger angle increasing significantly
the length of the drift paths as the electron would 
make a long spiral around the beam line before to reach the GEMs.

One of the important problematic to develop the radial TPC was to obtain a 
good support structure of the GEM foils, allowing an easy installation of the 
GEM in good conditions and with high precision. At the same time, we wanted to 
keep the material budget small in the forward region where we detect other 
particles in subsequent detectors. We successfuly realized these prerequisites 
by using fiber glass rings glued to each ends of the GEM foils to form self 
supporting cylinders that could be installed independently in the RTPC after 
all gluing and soldering operations.  Because the rigidity of the high 
rigidity of the GEM foil, the structure was selfsupporting and only the 
backward part of the cylinder was fixed to the main mechanical structure. This 
design only left a light fiberglass ring in the forward region, reducing to a 
minimum secondary interactions.

@@

The drift path is the trajectory followed by the electrons released through 
ionization in the gas. Software exists to calculate the drift paths, in particular
MAGBOLTZ~\cite{MAGBOLTZ}, but it requires knowledge of the detector's 
geometry, gas mixture composition, and of course the electric and magnetic 
fields over the whole volume of the detector. We used such calculation as a 
first calibration, but, 
as can be seen with the drift speed, gas composition is far from stable in the 
chamber. Moreover, the 4 $\mu$m foil used as a cathode is easily deformed, such
that we expect the geometry accuracy to be around the millimeter precision. 
These problems, already 
encountered for the BoNuS RTPC calibration~\cite{BONUS-NIM}, motivated the 
acquisition of specific calibration runs. These are taken with a lower energy 
electron beam (1.20 and 1.27 GeV) to enhance the cross section of the elastic 
process ($e^{4}$He$\rightarrow e^{4}$He). In this process, the measurement of
electron kinematics allows to calculate the Helium nucleus kinematic. It is
by comparing this calculated kinematic to the measured one that we fine tune
the drift paths independently of our knowledge of the exact conditions in the 
chamber.

@@

The second method is to compare the experimental ADCs to the energy deposited in 
GEANT4 by similar simulated tracks (using the same elastic events than for the drift 
paths calibration). This requires a very good GEANT4 simulation 
including drift paths, but also the spread of the charges along the path
before reaching the pad, so that the simulated hits match the experimental 
ones. Moreover, the simulation has to match the data acquisition (DAQ) features 
that can lead to cutting out hits. After setting the simulation properly, we 
compared simulation to experiment on an event by event basis as shown in 
figure~\ref{fig:EVENT_adc_tdc}. In this step, the gain for each pad is calculated 
as the ratio of the measured ADCs to the simulated deposited energy.  
Then, these gains are refined using correcting factors obtained from a 
sample of good tracks. For each track, we calculate the corrected 
energy deposit on a pad and compare it to the average
deposit recorded by the other pads. 
Results of this second method are shown in Figure~\ref{fig:dedx_p_exp_2nd},
from these we concluded that the gains of the 
second method match best the theoretical lines.
